---
title: "Linkage map and QTL mapping for _Ipomoea trifida_"
author: "Ana Paula Alves da Mata"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc_depth: 4
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 6, fig.height = 6, cache = TRUE, cache.lazy = FALSE, dev = "png", warning = FALSE)
```

# Experimental data and phenotypic analysis

### About the data

The data consists in **188 individuals**, including both parents, from a mapping population derived from the **M9×M19** cross (***Ipomoea trifida*** - 2*n* = 2*x* = 30) were evaluated. 

This population was phenotyped at the International Potato Center (CIP) in San Ramon, Peru, in 2016 under screen house condition, in a **randomized complete block design** with **four replications**.

```{r}
data2016 <- read.csv("M9xM19_screenhouse.csv", header = TRUE, sep = ",")
```

```{r}
colnames(data2016)
length(unique(data2016$CIPN)) #188 different individuals
str(data2016)
table(data2016$Rep) #4 replications with the 188 individuals
```

To better organize the data we'll analyze:

```{r, eval = -(8:9)}
data2016 <- data2016[,c(2,4,16:19,5,7,9,11,13,15,34)]

colnames(data2016) <- c("geno","rep","LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4",
                        "LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA")

data2016$geno <- as.factor(data2016$geno)
data2016$rep  <- as.factor(data2016$rep)

summary(data2016)
head(data2016)
```

## The mixed model

$$y_{ij} = \mu + b_j + g_i + e_{ij}$$

where $y_{ij}$ is the phenotype of the $i$th genotype in $j$th block; \\mu is the overall mean; $b_j$ is the fixed effect of the $j$th block (our replications); $g_i$ is the random effect of the $i$ genotype where $g_i \sim \mathcal{N}(0,\sigma_g^2)$ , and $\sigma_g^2$ is the genetic variance; and $e_{ij}$ is the random error, where $e_{ij} \sim \mathcal{N}(0, \sigma_e^2)$, and $\sigma_e^2)$ is the residual variance.

#### Broad-sense heritability

$$H^2 = \frac{\sigma_g^2}{\sigma_g^2 + \frac{\sigma_e^2}{r}}$$
where $r$ is the number of blocks (replications).

```{r, message = FALSE, warning = FALSE}
#install.packages("sommer")
library(sommer) #please make sure it's version 4.1.2
```

Here we have the analysis in a loop; it may take some time on some computers:

```{r, eval = FALSE}
mod.list <- modr.list <- LRT.list <- blup.list <- varcomp.list <- h2.list <- list()

for(trait in c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4","LEFTPP_TP1",
               "LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA")) {
  print(paste("Analysis for", trait))
  
  mod <- mmer(fixed = as.formula(paste(trait,"~ rep")),
              random = ~ geno,
              rcov = ~ units,
              data = data2016,
              date.warning = FALSE)
  mod.list[[trait]] <- mod
  
  #reduced model 
  modr <- mmer(fixed = as.formula(paste(trait,"~ rep")),
               rcov = ~ units,
               data = data2016,
               date.warning = FALSE)
  modr.list[[trait]] <- modr
  
  #likelihood ratio test (LRT)
  LRTmod <- anova(mod,modr)
  LRT.list[[trait]] <- LRTmod
  
  #BLUPS
  blups.mod <- predict(mod, classify = "geno")
  blup.list[[trait]] <- blups.mod
  
  #variance components
  var <- summary(mod)$varcomp
  varcomp.list[[trait]] <- var
  
  #heritability
  h2 <- vpredict(mod, h2 ~ V1/(V1+V2/4))
  h2.list[[trait]] <- h2
  
  rm(mod);rm(modr);rm(LRTmod);rm(blups.mod);rm(var);rm(h2)
}
```

Saving the results:

```{r, eval = FALSE}
save(mod.list, file = "mod.list.RData")
save(modr.list, file = "modr.list.RData")
save(LRT.list, file = "LRT.list.RData")
save(blup.list, file = "blup.list.RData")
save(varcomp.list, file = "varcomp.list.RData")
save(h2.list, file = "h2.list.RData")
```

Loading the results to avoid rerunning the entire analysis:

```{r}
load("mod.list.RData")
load("modr.list.RData")
load("LRT.list.RData")
load("blup.list.RData")
load("varcomp.list.RData")
load("h2.list.RData")
```

### Means and coefficients of variation

Here we see the mean (**m**) and the coefficients of variation (**cv**) of each trait:

```{r, eval = FALSE}
sd1 <- sqrt(varcomp.list[["LEAFSHAP1"]][["VarComp"]][[1]])
m1 <- mean(blup.list$LEAFSHAP1$pvals$predicted.value)
cv1 <- sd1/m1 
cv1

sd2 <- sqrt(varcomp.list[["LEAFSHAP2"]][["VarComp"]][[1]])
m2 <- mean(blup.list$LEAFSHAP2$pvals$predicted.value)
cv2 <- sd2/m2
cv2

sd3 <- sqrt(varcomp.list[["LEAFSHAP3"]][["VarComp"]][[1]])
m3 <- mean(blup.list$LEAFSHAP3$pvals$predicted.value)
cv3 <- sd3/m3 
cv3

sd4 <- sqrt(varcomp.list[["LEAFSHAP4"]][["VarComp"]][[1]])
m4 <- mean(blup.list$LEAFSHAP4$pvals$predicted.value)
cv4 <- sd4/m4 
cv4

sd5 <- sqrt(varcomp.list[["LEFTPP_TP1"]][["VarComp"]][[1]])
m5 <- mean(blup.list$LEFTPP_TP1$pvals$predicted.value)
cv5 <- sd5/m5 
cv5

sd6 <- sqrt(varcomp.list[["LEFTPP_TP2"]][["VarComp"]][[1]])
m6 <- mean(blup.list$LEFTPP_TP2$pvals$predicted.value)
cv6 <- sd6/m6 
cv6

sd7 <- sqrt(varcomp.list[["LEFTPP_TP3"]][["VarComp"]][[1]])
m7 <- mean(blup.list$LEFTPP_TP3$pvals$predicted.value)
cv7 <- sd7/m7
cv7

sd8 <- sqrt(varcomp.list[["PLANTH_TP1"]][["VarComp"]][[1]])
m8 <- mean(blup.list$PLANTH_TP1$pvals$predicted.value)
cv8 <- sd8/m8
cv8

sd9 <- sqrt(varcomp.list[["PLANTH_TP2"]][["VarComp"]][[1]])
m9 <- mean(blup.list$PLANTH_TP2$pvals$predicted.value)
cv9 <- sd9/m9 
cv9

sd10 <- sqrt(varcomp.list[["VINDIA"]][["VarComp"]][[1]])
m10 <- mean(blup.list$VINDIA$pvals$predicted.value)
cv10 <- sd10/m10 
cv10

sd11 <- sqrt(varcomp.list[["LA"]][["VarComp"]][[1]])
m11 <- mean(blup.list$LA$pvals$predicted.value)
cv11 <- sd11/m11
cv11
```

### Variance components and heritabilities

To make it easier, we extracted the variance components and the heritabilities of the traits into an XLSX file:

```{r}
varcomp <- cbind(t(varcomp.list[["LEAFSHAP1"]]), t(varcomp.list[["LEAFSHAP2"]]),
                 t(varcomp.list[["LEAFSHAP3"]]), t(varcomp.list[["LEAFSHAP4"]]),
                 t(varcomp.list[["LEFTPP_TP1"]]), t(varcomp.list[["LEFTPP_TP2"]]),
                 t(varcomp.list[["LEFTPP_TP3"]]), t(varcomp.list[["PLANTH_TP1"]]),
                 t(varcomp.list[["PLANTH_TP2"]]), t(varcomp.list[["VINDIA"]]),
                 t(varcomp.list[["LA"]]))

varcomp <- t(varcomp[1:2, ])
colnames(varcomp) <- c("Variance components","Standard error")
```

```{r}
h2 <- cbind(t(h2.list[["LEAFSHAP1"]]), t(h2.list[["LEAFSHAP2"]]), 
            t(h2.list[["LEAFSHAP3"]]), t(h2.list[["LEAFSHAP4"]]),
            t(h2.list[["LEFTPP_TP1"]]),  t(h2.list[["LEFTPP_TP2"]]),
            t(h2.list[["LEFTPP_TP3"]]), t(h2.list[["PLANTH_TP1"]]),
            t(h2.list[["PLANTH_TP2"]]), t(h2.list[["VINDIA"]]), 
            t(h2.list[["LA"]]))

colnames(h2) <- c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4",
                  "LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3",
                  "PLANTH_TP1","PLANTH_TP2","VINDIA","LA")

h2 <- t(h2[1:2, ])
colnames(h2) <- c("Estimate","Standard error")
```

```{r, eval = FALSE}
#install.packages("rJava")
#install.packages("writexl")
#install.packages("xlsx")
library(rJava)
library(writexl)
```

```{r, eval = FALSE}
xlsx::write.xlsx2(varcomp,paste0("GeneticParameters_2016.xlsx"), sheetName = "Variance components")
xlsx::write.xlsx2(h2,paste0("GeneticParameters_2016.xlsx"), sheetName = "Heritabilities", append = TRUE)
```

## Correlations

Pairwise Pearson’s correlations assuming a significant global level of 5%.

Loading the data:

```{r}
load("blup.list.RData")
```

```{r}
blups2016 <- data.frame(cbind(blup.list$LEAFSHAP1$pvals$predicted.value,
                              blup.list$LEAFSHAP2$pvals$predicted.value,
                              blup.list$LEAFSHAP3$pvals$predicted.value,
                              blup.list$LEAFSHAP4$pvals$predicted.value,
                              blup.list$LEFTPP_TP1$pvals$predicted.value,
                              blup.list$LEFTPP_TP2$pvals$predicted.value,
                              blup.list$LEFTPP_TP3$pvals$predicted.value,
                              blup.list$PLANTH_TP1$pvals$predicted.value,
                              blup.list$PLANTH_TP2$pvals$predicted.value,
                              blup.list$VINDIA$pvals$predicted.value,
                              blup.list$LA$pvals$predicted.value))

colnames(blups2016) <- c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4","LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA")

rownames(blups2016) <- blup.list$pvals$geno

cor(blups2016)
```

```{r, message = FALSE, warning = FALSE}
library(PerformanceAnalytics)

chart.Correlation(blups2016, histogram = TRUE, method = "pearson")

correl2016 <- cor(blups2016)
```

A function for plotting correlations:

```{r}
chart.Correlation.nostars <- function (R, histogram = TRUE, method = c("pearson", "kendall", "spearman"), ...)
{
  x = checkData(R, method = "matrix")
  if (missing(method))
    method = method[1]
  panel.cor <- function(x, y, digits = 2, prefix = "",
                        use = "pairwise.complete.obs", method = "pearson",
                        cex.cor, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y, use = use, method = method)
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste(prefix, txt, sep = "")
    if (missing(cex.cor))
      cex <- 0.8/strwidth(txt)
    test <- cor.test(as.numeric(x), as.numeric(y), method = method)
    text(0.5, 0.5, txt, cex = cex * (abs(r) + 0.3)/1.3)
  }
  
  f <- function(t) {
    dnorm(t, mean = mean(x), sd = sd.xts(x))
  }
  
  dotargs <- list(...)
  dotargs$method <- NULL
  rm(method)
  
  hist.panel = function(x, ... = NULL) {
    par(new = TRUE)
    hist(x, col = "light gray", probability = TRUE,
         axes = FALSE, main = "", breaks = "FD")
    lines(density(x, na.rm = TRUE), col = "red", lwd = 1)
    rug(x)
  }
  
  if (histogram)
    pairs(x, gap = 0, lower.panel = panel.smooth, 
          upper.panel = panel.cor,
          diag.panel = hist.panel)
  else 
    pairs(x, gap = 0, lower.panel = panel.smooth, 
          upper.panel = panel.cor)
}
```

```{r, fig.width = 14, fig.height = 8, message = FALSE, warning = FALSE}
library(ggplot2)

get_upper_tri <- function(correl2016) {
  correl2016[lower.tri(correl2016)] <- NA
  return(correl2016)
}
upper_tri <- get_upper_tri(correl2016)
```

Plot:

```{r, fig.width = 14, fig.height = 8}
library(reshape2)

cor_melt <- melt(upper_tri, na.rm = TRUE)
is.num <- sapply(cor_melt, is.numeric)
cor_melt[is.num] <- lapply(cor_melt[is.num], round, 2)

ggheatmap <- ggplot(cor_melt, aes(Var2, Var1, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "purple", high = "orange", mid = "white",  midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Pearson\nCorrelation") +
  scale_y_discrete (position = "right") +
  theme_minimal() + # minimal theme
  theme(axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45, size = 15), axis.text.y = element_text(size = 15)) +
  coord_fixed() +
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 5) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank(), panel.background = element_blank(), axis.ticks = element_blank(), legend.justification = c(1, 0), legend.position = c(0.4, 0.7), legend.direction = "horizontal") +
  guides(fill = guide_colorbar(barwidth = 15, barheight = 2, title.position = "top", title.hjust = 0.5)); ggheatmap

ggsave("ggheatmap2016.pdf", width = 18, height = 15)
```

# Genotypic analysis

## Genetic map

```{r, message = FALSE}
library(onemap)
library(tidyverse)
```

### Reading data and filtering

Load the data into Onemap:

```{r}
vcf <- onemap_read_vcfR(vcf = "all_variants_trifida_merged.vcf",
                        cross = "outcross", 
                        parent1 = "CIP107665.9.merged", 
                        parent2 = "CIP107665.19.merged",
                        only_biallelic = TRUE, 
                        verbose = TRUE)

save(vcf, file = "vcf.RData")
load("vcf.RData")
vcf
```

Plot of `r vcf$n.mar` SNPs vs. `r vcf$n.ind` individuals:

```{r}
plot(vcf)
```

Filter for *marker* missing data:

```{r, echo = -(2:3)}
vcf_filtered <- filter_missing(vcf, by = "markers", threshold = 0.20)
rm(vcf)
gc()

vcf_filtered
```

Plot of `r vcf_filtered$n.mar` filtered SNPs (markers with missing \< 20%):

```{r}
plot(vcf_filtered)
```

Filter for *individual* missing data:

```{r}
vcf_filtered <- filter_missing(vcf_filtered, by = "individuals", threshold = 0.50)
save(vcf_filtered, file = "vcf_filtered.RData")

write_onemap_raw(vcf_filtered, file.name = "vcf_filtered.raw")
vcf_filtered
```

Plot of `r vcf_filtered$n.ind` filtered individuals (markers with missing \< 50%):

```{r}
plot(vcf_filtered)
```

Perform segregation test:

```{r, echo = -(3:4)}
seg_test <- test_segregation(vcf_filtered)
save(seg_test, file = "seg_test.RData")

rm(vcf_filtered)
gc()

print(seg_test)
plot(seg_test)

no_dist <- select_segreg(seg_test, distorted = FALSE, numbers = TRUE)
length(no_dist)
save(no_dist, file = "no_dist.RData")
```

Plotting $\chi^2$ tests across chromosomes:

```{r, fig.width = 7, fig.height = 5}
chi_data <- print(seg_test)
head(chi_data)

chi_data[,c("Chr", "bp")] <- matrix(unlist(strsplit(chi_data$Marker, "_")), byrow = TRUE, ncol = 2)

colnames(chi_data)[c(3,4)] <- c("Chi", "P")
chi_data$bp <- as.numeric(chi_data$bp)

chi_data %>% 
  mutate(Distorted = ifelse(chi_data$P < 0.05/nrow(chi_data), "Yes", "No")) %>%
  filter(Chr != "Chr00") %>%
  ggplot() +
  geom_point(aes(x = bp/1000000, y = -log10(P), color = Distorted), shape = 1, size = 0.5, alpha = 0.5) +
  scale_color_manual(values = c("darkgrey", "red")) +
  geom_abline(intercept = -log10(0.05/nrow(chi_data)), slope = 0, linewidth = 0.2, color = "red") +
  facet_wrap(Chr ~ ., ncol = 3, strip.position = "right") + 
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40)) +
  labs(y = expression(-"log"[10](italic("P"))), x = "Genome position (Mbp)") +
  theme_bw() + 
  theme(legend.position = "bottom")

ggsave("segr_test_chr.pdf", width = 7, height = 5)
```

### Recombination frequency computation

Compute two-point recombination frequencies:

```{r}
twopts <- rf_2pts(input.obj = vcf_filtered)
save(twopts, file = "twopts.RData")

load("twopts.RData")
```

Filter `length(no_dist)` non-distorted markers:

```{r}
mark_no_dist <- make_seq(twopts, c(no_dist))
```

Visualize recombination frequency matrix for `length(mark_no_dist$seq.num)` markers:

```{r, fig.width = 15, fig.height = 15, dpi = 400}
rf_graph_table(mark_no_dist, main = paste(length(mark_no_dist$seq.num), "markers"), mrk.axis = "none")
```

### Grouping

#### Genome-based grouping

Group according to genome information:

```{r}
chr_map <- vector("list", 16)

names(chr_map) <- paste0("Chr", sprintf("%02d", (1:16)-1))
for(c in 1:16) {
  chr_map[[c]] <- make_seq(twopts, paste0("Chr", sprintf("%02d", c-1)))
  chr_map[[c]] <- onemap::map(chr_map[[c]], global_error = 0.15, tol = 1e-03)
  if(c > 1) {
    chr_map[[c]]$data.name <- NA
    chr_map[[c]]$twopt <- NA
  }
}

save(chr_map, file = "chr_map.RData")
```

Visualize heatmaps:

```{r, echo=-1}
load("chr_map.RData")
for(c in 2:16) {
  chr_map[[c]]$data.name <- chr_map[[1]]$data.name
  chr_map[[c]]$twopt <- chr_map[[1]]$twopt
}

for(c in 2:16) {
  print(paste0("Chr", sprintf("%02d", c-1)))
  print(chr_map[[c]])
  which_chr <- table(chr_map[[c]]$data.name$CHROM[chr_map[[c]]$seq.num])
  which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); print(which_chr)
  print(rf_graph_table(chr_map[[c]], main = which_chr, mrk.axis = "none"))
}
```

Visualize maps:

```{r}
draw_map(chr_map, names = FALSE, grid = FALSE, cex.mrk = 0.7)
```

Linkage groups *versus* chromosomes:

```{r}
lgs_vs_chr <- data.frame(row.names = c("ID", "Chr", "bp", "Lgr", "cM"))
for(c in 2:16) {
  ID <- chr_map[[c]]$seq.num
  Chr <- chr_map[[c]]$data.name$CHROM[chr_map[[c]]$seq.num]
  bp <- chr_map[[c]]$data.name$POS[chr_map[[c]]$seq.num]
  Lgr <- rep(names(chr_map)[c], length(chr_map[[c]]$seq.num))
  cM <- c(0, cumsum(kosambi(chr_map[[c]]$seq.rf)))
  lgs_vs_chr <- rbind(lgs_vs_chr, data.frame(ID, Chr, bp, Lgr, cM))
}

head(lgs_vs_chr)
```

```{r, fig.height = 8, fig.width = 13}
ggplot(lgs_vs_chr) +
  geom_point(aes(x = cM, y = bp/1000000, color = Chr), shape = 1, alpha = 0.5) + 
  facet_grid(Chr ~ Lgr, scales = "free", space = "free") + 
  scale_x_continuous(breaks = seq(0, round(max(lgs_vs_chr$cM), digits = -2), by = round(max(lgs_vs_chr$cM), digits = -2)/5)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40)) +
  labs(x = "Map position (cM)", y = "Genome position (Mbp)") +
  theme_bw() + 
  theme(panel.spacing=unit(0, "lines"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave("lgs_vs_chr_genome.pdf", height = 15, width = 30)
```

Map summary:

```{r, echo = -(13:14)}
map_summary <- data.frame(row.names = c("Chr", "N.mrk", "N.distorted", "N.nondistorted", "Lenght", "Max.gap"))

for(c in 1:15) {
  Chr <- sprintf("%02d", c)
  N.mrk <- length(chr_map[[c]]$seq.num)
  Length <- tail(cumsum(kosambi(chr_map[[c]]$seq.rf)), 1)
  N.distorted <- sum(!chr_map[[c]]$seq.num %in% no_dist)
  N.nondistorted <- sum(chr_map[[c]]$seq.num %in% no_dist)
  Max.gap <- max(kosambi(chr_map[[c]]$seq.rf))
  map_summary <- rbind(map_summary, data.frame(Chr, N.mrk, Length = round(Length, 2), N.distorted, N.nondistorted, Max.gap = round(Max.gap, 2)))
}

map_summary <- rbind(map_summary, data.frame(Chr = "Total", N.mrk = sum(map_summary$N.mrk), 
                                             N.distorted = sum(map_summary$N.distorted), 
                                             N.nondistorted = sum(map_summary$N.nondistorted), 
                                             Length = sum(map_summary$Length),
                                             Max.gap = sum(map_summary$Max.gap)))

knitr::kable(map_summary)

rm(chr_map)
gc()
```

#### *De novo* grouping based on UPGMA

Group according to recombination frequency estimates (*de novo*) using UPGMA method:

```{r, eval = -(1:4), echo = -5, fig.width = 8, dpi = 400}
lgs <- group_upgma(mark_no_dist, expected.groups = 16, inter = FALSE, comp.mat = TRUE)

lgs$groups[lgs$groups == 2] <- 1
lgs$groups <- ifelse(test = lgs$groups>1, yes = lgs$groups-1, no = lgs$groups)

save(lgs, file = "lgs_upgma.RData")

load("lgs_upgma.RData")

lgs$seq.vs.grouped.snp
sort(table(lgs$groups))
plot(lgs)
lgs$n.groups <- length(table(lgs$groups))
```

```{r, eval = FALSE}
lgs_map <- only_main_chr <- markers_to_try <- vector("list", lgs$n.groups)

names(lgs_map) <- names(only_main_chr) <- names(markers_to_try) <- paste0("Lgr", sprintf("%02d", 1:lgs$n.groups))

for(c in 1:lgs$n.groups) {
  print(paste0("Lgr", sprintf("%02d", c)))
  lgs_map[[c]] <- make_seq(lgs, c)
  temp_chrom <- lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num]
  only_main_chr[[c]] <- lgs_map[[c]]$seq.num[which(temp_chrom == names(which.max(table(temp_chrom))))]
  markers_to_try[[c]] <- lgs_map[[c]]$seq.num[which(temp_chrom != names(which.max(table(temp_chrom))))]
  lgs_map[[c]] <- make_seq(lgs$twopt, sort(only_main_chr[[c]]))
  lgs_map[[c]] <- onemap::map(lgs_map[[c]], global_error = 0.15, tol = 1e-03)
  if(c > 1) {
    lgs_map[[c]]$data.name <- NA
    lgs_map[[c]]$twopt <- NA
  } 
}

save(lgs_map, file = "lgs_map_upgma.RData")
save(only_main_chr, markers_to_try, file = "markers_upgma.RData")
```

Visualize heatmaps:

```{r, echo = -(1:2)}
load("lgs_map_upgma.RData")
load("markers_upgma.RData")
for(c in 2:15) {
  lgs_map[[c]]$data.name <- lgs_map[[1]]$data.name
  lgs_map[[c]]$twopt <- lgs_map[[1]]$twopt
}

for(c in 1:15) {
  print(paste0("Lgr", sprintf("%02d", c)))
  print(lgs_map[[c]])
  which_chr <- table(lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num])
  which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", ")
  print(rf_graph_table(lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none"))
}
```

Visualize maps:

```{r}
draw_map(lgs_map, names = FALSE, grid = FALSE, cex.mrk = 0.7)
```

Linkage groups *versus* chromosomes:

```{r}
lgs_vs_chr <- data.frame(row.names = c("ID", "Chr", "bp", "Lgr", "cM"))

for(c in 1:lgs$n.groups) {
  ID <- lgs_map[[c]]$seq.num
  Chr <- lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num]
  bp <- lgs_map[[c]]$data.name$POS[lgs_map[[c]]$seq.num]
  Lgr <- rep(names(lgs_map)[c], length(lgs_map[[c]]$seq.num))
  cM <- c(0, cumsum(kosambi(lgs_map[[c]]$seq.rf)))
  lgs_vs_chr <- rbind(lgs_vs_chr, data.frame(ID, Chr, bp, Lgr, cM))
}

head(lgs_vs_chr)
```

```{r, fig.height = 8, fig.width = 13}
ggplot(lgs_vs_chr) +
  geom_point(aes(x = cM, y = bp/1000000, color = Chr), shape = 1, alpha = 0.5) + 
  facet_grid(Chr ~ Lgr, scales = "free", space = "free") + 
  scale_x_continuous(breaks = seq(0, round(max(lgs_vs_chr$cM), digits = -2), by = round(max(lgs_vs_chr$cM), digits = -2)/5)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40)) +
  labs(x = "Map position (cM)", y = "Genome position (Mbp)") +
  theme_bw() + 
  theme(panel.spacing=unit(0, "lines"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave("lgs_vs_chr_upgma.pdf", height = 8, width = 13)
```

Map summary:

```{r}
map_summary <- data.frame(row.names = c("Lgr", "Chr", "N.mrk", "Lenght", "Max.gap"))

all_Chr <- c()

for(c in 1:lgs$n.groups) {
  Lgr <- sprintf("%02d", c)
  which_mrk <- lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num]
  all_Chr <- c(all_Chr, which_mrk)
  Chr <- paste(names(table(which_mrk)), table(which_mrk), sep = "=", collapse = ", ")
  N.mrk <- length(lgs_map[[c]]$seq.num)
  Length <- tail(cumsum(kosambi(lgs_map[[c]]$seq.rf)), 1)
  Max.gap <- max(kosambi(lgs_map[[c]]$seq.rf))
  map_summary <- rbind(map_summary, data.frame(Lgr, Chr, N.mrk, Length = round(Length, 2), Max.gap = round(Max.gap, 2)))
}

all_Chr <- paste(names(table(all_Chr)), table(all_Chr), sep = "=", collapse = ", ")

map_summary <- rbind(map_summary, data.frame(Lgr = "Total", Chr = all_Chr, 
                                             N.mrk = sum(map_summary$N.mrk), 
                                             Length = sum(map_summary$Length),
                                             Max.gap = sum(map_summary$Max.gap)))

knitr::kable(map_summary)
```

#### Split groups based on gaps

Estimate number of expected groups based on gaps from genome information:

```{r}
gap <- 5 #gap size that split chr
gr <- 0
groups <- c()

for(c in 1:lgs$n.groups) {
  mrk_gap <- 0
  temp.groups <- c()
  temp.gap <- which(kosambi(lgs_map[[c]]$seq.rf) > gap)
  if(length(temp.gap) > 0) {
    mrk_gap <- c(mrk_gap, temp.gap)
  } 
  
  mrk_gap <- c(mrk_gap, length(lgs_map[[c]]$seq.num))
  for(m in 1:(length(mrk_gap)-1)) {
    gr <- gr + 1
    temp.groups <- c(temp.groups, rep(gr, mrk_gap[m+1]-mrk_gap[m]))
  }
  
  attr(temp.groups, "names") <- as.character(lgs_map[[c]]$seq.num)
  groups <- c(groups, temp.groups)
}

table(groups)
```

We replace the `lgs` object info with the one from gap-assisted grouping:

```{r}
lgs$n.groups <- length(table(groups))
no.groups <- lgs$seq.num[!lgs$seq.num %in% sort(as.numeric(names(groups)))]
na.groups <- rep(NA, length(no.groups))
attr(na.groups, "names") <- as.character(no.groups)

groups <- c(na.groups, groups)
groups <- groups[order(as.numeric(names(groups)))]
lgs$groups <- groups

count_chr <- lgs_map <- vector("list", lgs$n.groups)
names(lgs_map) <- paste0("Lgr", sprintf("%02d", 1:lgs$n.groups))
n_min <- 2 #minimum number of markers per group
first_chr <- c()

for(c in 1:lgs$n.groups) {
  print(paste0("Lgr", sprintf("%02d", c)))
  lgs_map[[c]] <- make_seq(lgs, c)
  count_chr[[c]] <- table(lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num])
  if(length(lgs_map[[c]]$seq.num) >= n_min) { 
    if(length(first_chr) == 0) first_chr <- c
    lgs_map[[c]] <- onemap::map(lgs_map[[c]], global_error = 0.15, tol = 1e-03)
  } 
  if(c > first_chr) {
    lgs_map[[c]]$data.name <- NA
    lgs_map[[c]]$twopt <- NA
  }
}

unlist(count_chr)

save(lgs, file = "lgs_gaps.RData")
save(lgs_map, file = "lgs_map_gaps.RData")
```

Visualize heatmaps:

```{r, echo = -(1:2)}
load("lgs_gaps.RData")
load("lgs_map_gaps.RData")

which(sapply(lgs_map, is.null) == FALSE) 
first_chr <- which(sapply(lgs_map, is.null) == FALSE)[1]
for(c in 1:length(lgs_map)) {
  lgs_map[[c]]$data.name <- lgs_map[[first_chr]]$data.name
  lgs_map[[c]]$twopt <- lgs_map[[first_chr]]$twopt
}

estimated <- c()
for(c in 1:length(lgs_map)) {
  if(length(lgs_map[[c]]$seq.phases) > 1){
    print(paste0("Lgr", sprintf("%02d", c)))
    estimated <- c(estimated, c)
    print(lgs_map[[c]])
    which_chr <- table(lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num])
    which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", ")
    print(rf_graph_table(lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none"))
  }
}
```

Visualize maps:

```{r}
draw_map(lgs_map[estimated], names = FALSE, grid = FALSE, cex.mrk = 0.7)
```

Linkage groups *versus* chromosomes:

```{r}
lgs_vs_chr <- data.frame(row.names = c("ID", "Chr", "bp", "Lgr", "cM")) 

for(c in estimated) {
  ID <- lgs_map[[c]]$seq.num
  Chr <- lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num]
  bp <- lgs_map[[c]]$data.name$POS[lgs_map[[c]]$seq.num]
  Lgr <- rep(names(lgs_map)[c], length(lgs_map[[c]]$seq.num))
  cM <- c(0, cumsum(kosambi(lgs_map[[c]]$seq.rf)))
  lgs_vs_chr <- rbind(lgs_vs_chr, data.frame(ID, Chr, bp, Lgr, cM))
}

head(lgs_vs_chr)
```

```{r, fig.height = 8, fig.width = 13}
ggplot(lgs_vs_chr) + 
  geom_point(aes(x = cM, y = bp/1000000, color = Chr), shape = 1, alpha = 0.5) + 
  facet_grid(Chr ~ Lgr, scales = "free", space = "free") + 
  scale_x_continuous(breaks = seq(0, round(max(lgs_vs_chr$cM), digits = -2), by = round(max(lgs_vs_chr$cM), digits = -2)/5)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40)) +
  labs(x = "Map position (cM)", y = "Genome position (Mbp)") +
  theme_bw() + 
  theme(panel.spacing = unit(0, "lines"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave("lgs_vs_chr_gaps.pdf", height = 8, width = 13)
```

Map summary:

```{r}
map_summary <- data.frame(row.names = c("Lgr", "Chr", "N.mrk", "Lenght", "Max.gap")) 

all_Chr <- c()

for(c in estimated) {
  Lgr <- sprintf("%02d", c)
  which_mrk <- lgs_map[[c]]$data.name$CHROM[lgs_map[[c]]$seq.num]
  all_Chr <- c(all_Chr, which_mrk)
  Chr <- paste(names(table(which_mrk)), table(which_mrk), sep = "=", collapse = ", ")
  N.mrk <- length(lgs_map[[c]]$seq.num)
  Length <- tail(cumsum(kosambi(lgs_map[[c]]$seq.rf)), 1)
  Max.gap <- max(kosambi(lgs_map[[c]]$seq.rf))
  map_summary <- rbind(map_summary, data.frame(Lgr, Chr, N.mrk, Length = round(Length, 2), Max.gap = round(Max.gap, 2)))
  map_summary <- rbind(map_summary, data.frame(Lgr, Chr, N.mrk, Length = 0, Max.gap = 0))
}

all_Chr <- paste(names(table(all_Chr)), table(all_Chr), sep = "=", collapse = ", ")

map_summary <- rbind(map_summary, data.frame(Lgr = "Total", Chr = all_Chr, 
                                             N.mrk = sum(map_summary$N.mrk), 
                                             Length = sum(map_summary$Length),
                                             Max.gap = sum(map_summary$Max.gap)))

map_summary <- rbind(map_summary, data.frame(Lgr = "Total", Chr = all_Chr, 
                                             N.mrk = sum(map_summary$N.mrk), 
                                             Length = 0, Max.gap = 0))

knitr::kable(map_summary)
```

#### Combine genome-based and recombination frequency-based information

As we identified misassemblies in the genome-based map, it's time to reorder them:

```{r, eval = FALSE}
chr_lgs_map <- vector("list", 15)
names(chr_lgs_map) <- paste0("Lgr", sprintf("%02d", 1:15))
unlist(count_chr)
```

```{r, eval = FALSE}
c <- 1

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[1]]$seq.num, lgs_map[[3]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 2

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[7]]$seq.num, rev(lgs_map[[6]]$seq.num)))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 3

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[8]]$seq.num, rev(lgs_map[[9]]$seq.num)))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 4

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[11]]$seq.num, lgs_map[[12]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 5

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[13]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 6

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[15]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 7

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[18]]$seq.num, rev(lgs_map[[20]]$seq.num), lgs_map[[19]]$seq.num[1:23], lgs_map[[21]]$seq.num, lgs_map[[22]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 8

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[25]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 9

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[14]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 10

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[17]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 11

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[4]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 12

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[27]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 13

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[5]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 14

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[26]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

```{r, eval = FALSE}
c <- 15

chr_lgs_map[[c]] <- make_seq(lgs_map[[c]]$twopt, c(lgs_map[[28]]$seq.num, lgs_map[[30]]$seq.num))
chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", "); which_chr

rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none")
chr_lgs_map[[c]]
```

Saving new linkage groups:

```{r, eval = FALSE}
for(c in 2:15) {
  chr_lgs_map[[c]]$data.name <- NA
  chr_lgs_map[[c]]$twopt <- NA
}

save(chr_lgs_map, file = "chr_lgs_map_fixed.RData")
```

Visualize heatmaps::

```{r, echo = -1}
load("chr_lgs_map_fixed.RData")
for(c in 2:15) {
  chr_lgs_map[[c]]$data.name <- chr_lgs_map[[1]]$data.name
  chr_lgs_map[[c]]$twopt <- chr_lgs_map[[1]]$twopt
}

for(c in 1:15) {
  print(paste0("Lgr", sprintf("%02d", c)))
  print(chr_lgs_map[[c]])
  which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
  which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", ")
  print(rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none"))
}
```

Visualize maps:

```{r}
draw_map(chr_lgs_map, names = FALSE, grid = FALSE, cex.mrk = 0.7)
```

Linkage groups *versus* chromosomes:

```{r}
lgs_vs_chr <- data.frame(row.names = c("ID", "Chr", "bp", "Lgr", "cM"))

for(c in 1:15) {
  ID <- chr_lgs_map[[c]]$seq.num
  Chr <- chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num]
  bp <- chr_lgs_map[[c]]$data.name$POS[chr_lgs_map[[c]]$seq.num]
  Lgr <- rep(names(chr_lgs_map)[c], length(chr_lgs_map[[c]]$seq.num))
  cM <- c(0, cumsum(kosambi(chr_lgs_map[[c]]$seq.rf)))
  lgs_vs_chr <- rbind(lgs_vs_chr, data.frame(ID, Chr, bp, Lgr, cM))
}

head(lgs_vs_chr)
```

```{r, fig.height = 8, fig.width = 13}
ggplot(lgs_vs_chr) +
  geom_point(aes(x = cM, y = bp/1000000, color = Chr), shape = 1, alpha = 0.5) + 
  facet_grid(Chr ~ Lgr, scales = "free", space = "free") + 
  scale_x_continuous(breaks = seq(0, round(max(lgs_vs_chr$cM), digits = -2), by = round(max(lgs_vs_chr$cM), digits = -2)/5)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40)) +
  labs(x = "Map position (cM)", y = "Genome position (Mbp)") +
  theme_bw() + 
  theme(panel.spacing=unit(0, "lines"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave("lgs_vs_chr_fixed.pdf", height = 8, width = 13)
```

Map summary:

```{r}
map_summary <- data.frame(row.names = c("Lgr", "Chr", "N.mrk", "Lenght", "Max.gap"))

Chrtotal <- c()

for(c in 1:15) {
  Lgr <- sprintf("%02d", c)
  which_mrk <- chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num]
  Chrtotal <- c(Chrtotal, which_mrk)
  Chr <- paste(names(table(which_mrk)), table(which_mrk), sep = "=", collapse = ", ")
  N.mrk <- length(chr_lgs_map[[c]]$seq.num)
  Length <- tail(cumsum(kosambi(chr_lgs_map[[c]]$seq.rf)), 1)
  Max.gap <- max(kosambi(chr_lgs_map[[c]]$seq.rf))
  map_summary <- rbind(map_summary, data.frame(Lgr, Chr, N.mrk, Length = round(Length, 2), Max.gap = round(Max.gap, 2)))
}

Chrtotal <- paste(names(table(Chrtotal)), table(Chrtotal), sep = "=", collapse = ", ")

map_summary <- rbind(map_summary, data.frame(Lgr = "Total", Chr = Chrtotal, 
                                             N.mrk = sum(map_summary$N.mrk), 
                                             Length = sum(map_summary$Length),
                                             Max.gap = sum(map_summary$Max.gap)))

knitr::kable(map_summary)
```

#### Try markers

Some groups have markers originally assigned to other chromosomes detected using UPGMA. The markers from the same chromosome that were not added in the previous step are also tried here. We try best positions for them:

```{r}
markers_to_try <- markers_to_try[c(1,5,6,7,8,9,11,12,3,10,2,14,4,13,15)]
names(markers_to_try) <- paste0("Lgr", sprintf("%02d", 1:15))
markers_to_try

markers_to_try[[1]] <- c(markers_to_try[[1]], lgs_map[[2]]$seq.num)
markers_to_try[[4]] <- c(markers_to_try[[4]], lgs_map[[10]]$seq.num)
markers_to_try[[6]] <- c(markers_to_try[[6]], lgs_map[[16]]$seq.num)
markers_to_try[[7]] <- c(markers_to_try[[7]], lgs_map[[19]]$seq.num[24:26])
markers_to_try[[8]] <- c(markers_to_try[[8]], lgs_map[[23]]$seq.num, lgs_map[[24]]$seq.num)

markers_to_try[[15]] <- c(markers_to_try[[15]], lgs_map[[29]]$seq.num)
fix_lgs <- which(sapply(markers_to_try, length) > 0)
fix_lgs

markers_to_try
```

Fixing linkage groups:

```{r, eval = FALSE}
for (c in fix_lgs) {
  print(paste0("Chr", sprintf("%02d", c)))
  try_res <- vector("list", length = length(markers_to_try[[c]]))
  for(m in 1:length(markers_to_try[[c]])) {
    print(paste("Try", markers_to_try[[c]][m]))
    try_temp <- try_seq(chr_lgs_map[[c]], markers_to_try[[c]][m])
    if(which.max(try_temp$LOD) %in% c(1,length(try_temp$LOD))) {
      try_res[m] <- list(NULL)
    } else {
      try_res[[m]] <- try_temp$try.ord[which.max(try_temp$LOD),]
    }
  }
  try_res <- try_res[!sapply(try_res, is.null)]
  if(length(try_res) > 0) {
    chr_lgs_map[[c]] <- make_seq(chr_lgs_map[[c]]$twopt, unique(c(t(sapply(try_res, cbind)))))
    chr_lgs_map[[c]] <- onemap::map(chr_lgs_map[[c]], global_error = 0.15, tol = 1e-03)
  } 
}

for(c in 2:15) {
  chr_lgs_map[[c]]$data.name <- NA
  chr_lgs_map[[c]]$twopt <- NA
}

save(chr_lgs_map, file = "chr_lgs_map_try.RData")
```

Visualize heatmaps:

```{r, echo = -1}
load("chr_lgs_map_try.RData")
for(c in 2:15) {
  chr_lgs_map[[c]]$data.name <- chr_lgs_map[[1]]$data.name
  chr_lgs_map[[c]]$twopt <- chr_lgs_map[[1]]$twopt
}

for(c in 1:15) {
  print(paste0("Lgr", sprintf("%02d", c)))
  print(chr_lgs_map[[c]])
  which_chr <- table(chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num])
  which_chr <- paste(names(which_chr), which_chr, sep = "=", collapse = ", ")
  print(rf_graph_table(chr_lgs_map[[c]], main = paste0("Lgr", sprintf("%02d", c), " (", which_chr, ")"), mrk.axis = "none"))
}
```

Visualize maps:

```{r}
draw_map(chr_lgs_map, names = FALSE, grid = FALSE, cex.mrk = 0.7)
```

Linkage groups *versus* chromosomes:

```{r}
lgs_vs_chr <- data.frame(row.names = c("ID", "Chr", "bp", "Lgr", "cM"))

for(c in 1:15) {
  ID <- chr_lgs_map[[c]]$seq.num
  Chr <- chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num]
  bp <- chr_lgs_map[[c]]$data.name$POS[chr_lgs_map[[c]]$seq.num]
  Lgr <- rep(names(chr_lgs_map)[c], length(chr_lgs_map[[c]]$seq.num))
  cM <- c(0, cumsum(kosambi(chr_lgs_map[[c]]$seq.rf)))
  lgs_vs_chr <- rbind(lgs_vs_chr, data.frame(ID, Chr, bp, Lgr, cM))
}

head(lgs_vs_chr)
```

```{r, fig.height = 8, fig.width = 13}
ggplot(lgs_vs_chr) +
  geom_point(aes(x = cM, y = bp/1000000, color = Chr), shape = 1, alpha = 0.5) + 
  facet_grid(Chr ~ Lgr, scales = "free", space = "free") + 
  scale_x_continuous(breaks = seq(0, round(max(lgs_vs_chr$cM), digits = -2), by = round(max(lgs_vs_chr$cM), digits = -2)/5)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50)) +
  labs(x = "Map position (cM)", y = "Genome position (Mbp)") +
  theme_bw() + 
  theme(panel.spacing=unit(0, "lines"), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave("lgs_vs_chr_try.pdf", height = 8, width = 13)
```

Map summary:

```{r}
map_summary <- data.frame(row.names = c("Lgr", "Chr", "N.mrk", "Lenght", "Max.gap"))
Chrtotal <- c()

for(c in 1:15) {
  Lgr <- sprintf("%02d", c)
  which_mrk <- chr_lgs_map[[c]]$data.name$CHROM[chr_lgs_map[[c]]$seq.num]
  Chrtotal <- c(Chrtotal, which_mrk)
  Chr <- paste(names(table(which_mrk)), table(which_mrk), sep = "=", collapse = ", ")
  N.mrk <- length(chr_lgs_map[[c]]$seq.num)
  Length <- tail(cumsum(kosambi(chr_lgs_map[[c]]$seq.rf)), 1)
  Max.gap <- max(kosambi(chr_lgs_map[[c]]$seq.rf))
  map_summary <- rbind(map_summary, data.frame(Lgr, Chr, N.mrk, Length = round(Length, 2), Max.gap = round(Max.gap, 2)))
}

Chrtotal <- paste(names(table(Chrtotal)), table(Chrtotal), 
                  sep = "=", collapse = ", ")

map_summary <- rbind(map_summary, data.frame(Lgr = "Total", Chr = Chrtotal, 
                                             N.mrk = sum(map_summary$N.mrk), 
                                             Length = sum(map_summary$Length),
                                             Max.gap = sum(map_summary$Max.gap)))

knitr::kable(map_summary)
```

## QTL analysis

Required packages:
```{r}
library(onemap)
library(tidyverse)
library(tidyr)
library(fullsibQTL)
library(ggplot2)
```

```{r}
load("blup.list.RData")

blups16 <- data.frame(blup.list$LEAFSHAP1$pvals$predicted.value,
                      blup.list$LEAFSHAP2$pvals$predicted.value,
                      blup.list$LEAFSHAP3$pvals$predicted.value,
                      blup.list$LEAFSHAP4$pvals$predicted.value,
                      blup.list$LEFTPP_TP1$pvals$predicted.value,
                      blup.list$LEFTPP_TP2$pvals$predicted.value,
                      blup.list$LEFTPP_TP3$pvals$predicted.value,
                      blup.list$PLANTH_TP1$pvals$predicted.value,
                      blup.list$PLANTH_TP2$pvals$predicted.value,
                      blup.list$VINDIA$pvals$predicted.value,
                      blup.list$LA$pvals$predicted.value)

colnames(blups16) <- c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4","LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA")

rownames(blups16) <- blup.list$LEAFSHAP1$pvals$geno

write.csv(blups16, file = "pheno16.csv", row.names = TRUE)
```

```{r}
#load("blup.modelo.list.RData")

#blups17 <- data.frame(blup.modelo.list$SRDW$pvals$predicted.value,
#                      blup.modelo.list$SRFW$pvals$predicted.value,
#                      blup.modelo.list$VDW$pvals$predicted.value,
#                      blup.modelo.list$VFW$pvals$predicted.value,
#                      blup.modelo.list$LFSIZE1$pvals$predicted.value,
#                      blup.modelo.list$LFSIZE2$pvals$predicted.value,
#                      blup.modelo.list$PETIOL1$pvals$predicted.value,
#                      blup.modelo.list$PETIOL2$pvals$predicted.value,
#                      blup.modelo.list$LAI$pvals$predicted.value,
#                      blup.modelo.list$NOPH$pvals$predicted.value,
#                      blup.modelo.list$VFW.plant$pvals$predicted.value,
#                      blup.modelo.list$VDW.plant$pvals$predicted.value,
#                      blup.modelo.list$VFW_sab$pvals$predicted.value,
#                      blup.modelo.list$VDW_sam$pvals$predicted.value,
#                      blup.modelo.list$SRFW_sam$pvals$predicted.value,
#                      blup.modelo.list$SRDW_sam$pvals$predicted.value,
#                      blup.modelo.list$SRFW.plant$pvals$predicted.value,
#                      blup.modelo.list$SRDW.plant$pvals$predicted.value,
#                      blup.modelo.list$NOSR$pvals$predicted.value,
#                      blup.modelo.list$NOSR.plant$pvals$predicted.value,
#                      blup.modelo.list$PTRN$pvals$predicted.value,
#                      blup.modelo.list$PTRN.plant$pvals$predicted.value,
#                      blup.modelo.list$PTRFW$pvals$predicted.value,
#                      blup.modelo.list$PTRFW.plant$pvals$predicted.value)

#colnames(blups17) <- c("SRDW","SRFW","VDW","VFW","LFSIZE1","LFSIZE2","PETIOL1","PETIOL2","LAI","NOPH","VFW.plant","VDW.plant","VFW_sab","VDW_sam","SRFW_sam","SRDW_sam","SRFW.plant","SRDW.plant","NOSR","NOSR.plant","PTRN","PTRN.plant","PTRFW","PTRFW.plant")
                       
#rownames(blups17) <- blup.modelo.list$SRDW$pvals$geno
            
#write.csv(blups17, file = "pheno17.csv", row.names = TRUE)
```

Plot - boxplot and violin:

```{r}
box16 <- data.frame(blup.list$LEAFSHAP1$pvals$predicted.value,
                    blup.list$LEAFSHAP2$pvals$predicted.value,
                    blup.list$LEAFSHAP3$pvals$predicted.value,
                    blup.list$LEAFSHAP4$pvals$predicted.value,
                    blup.list$LEFTPP_TP1$pvals$predicted.value,
                    blup.list$LEFTPP_TP2$pvals$predicted.value,
                    blup.list$LEFTPP_TP3$pvals$predicted.value,
                    blup.list$PLANTH_TP1$pvals$predicted.value,
                    blup.list$PLANTH_TP2$pvals$predicted.value,
                    blup.list$VINDIA$pvals$predicted.value,
                    blup.list$LA$pvals$predicted.value)

colnames(box16) <- c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4","LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA")

box16$geno <- blup.list$LEAFSHAP1$pvals$geno

year <- c("2016")
blups16$Year <- year
```

```{r}
box16_long <- box16 %>% pivot_longer(cols = c("LEAFSHAP1","LEAFSHAP2","LEAFSHAP3","LEAFSHAP4","LEFTPP_TP1","LEFTPP_TP2","LEFTPP_TP3","PLANTH_TP1","PLANTH_TP2","VINDIA","LA"), names_to = "Phenotype", values_to = "Value")

box16_long$Year <- "2016"

box16_long$Phenotype <- factor(box16_long$Phenotype, levels = unique(box16_long$Phenotype))
box16_long$geno <- gsub("CIP107665.19", "M19", box16_long$geno)
box16_long$geno <- gsub("CIP107665.9", "M9", box16_long$geno)
head(box16_long)
```

```{r}
ggplot(box16_long, aes(x = Year, y = Value)) + 
  geom_violin(aes(group = Year), color = "salmon") + 
  geom_point(data = box16_long %>% filter(geno %in% c("M19", "M9")), 
             aes(shape = geno), 
             color = "black", alpha = 0.7, size = 1, stroke = 1.5) + 
  scale_shape_manual(values = c(3,4)) + 
  facet_wrap(. ~ Phenotype, scales = "free_y", nrow = 3) + 
  labs(y = "Predict Value", shape = "Parent") + 
  theme_bw() + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        legend.position = "right", 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 10))

ggsave("violin.png", width = 8, height = 6, units = "in", dpi = 300)
```

### Reading data and filtering

```{r}
inds <- scan("vcf_filtered.raw", what = "character", skip = 2, nlines = 1)
inds
```

BLUPs:

```{r}
pheno16 <- read.csv("pheno16.csv", header = TRUE, row.names = 1)
colnames(pheno16) <- paste(colnames(pheno16), "2016", sep = "_")

for(p in 1:ncol(pheno16)) {
  write.table(paste(paste0("*", colnames(pheno16)[p]), 
                    paste(pheno16[inds, p], collapse = " "), collapse = "\n", sep = " "), 
              file = "vcf_filtered.raw", row.names = FALSE, 
              col.names = FALSE, quote = FALSE, append = TRUE)
}

#pheno17 <- read.csv("pheno17.csv", header = TRUE, row.names = 1)
#colnames(pheno17) <- paste(colnames(pheno17), "2017", sep = "_")
#for(p in 1:ncol(pheno17)) {
#  write.table(paste(paste0("*", colnames(pheno17)[p]), paste(pheno17[inds,p], collapse = " "), collapse = "\n", sep = " "), file = "vcf_filtered.raw", row.names = FALSE, col.names = FALSE, quote = FALSE, append = TRUE)
#}
```

Loading map:

```{r, echo = -1}
load("chr_lgs_map_try.RData")
for(c in 2:15) {
  chr_lgs_map[[c]]$data.name <- chr_lgs_map[[1]]$data.name
  chr_lgs_map[[c]]$twopt <- chr_lgs_map[[1]]$twopt
}
```

```{r}
#Raw data
fs_data <- read_onemap(inputfile = "vcf_filtered.raw")
fs_data

#Information of linkage groups, step and mapping function
fsib <- create_fullsib(fs_data, map.list = chr_lgs_map, step = 1, map.function = "kosambi")
fsib
```

### QTL mapping - CIM

#### Cofactors

```{r}
#`%||%` <- function (x, y) {if (is_null(x)) y else x}
```

```{r, eval = FALSE}
cofs_fs <- vector("list", ncol(fsib$pheno))
pdf("cofs.pdf", width = 16, height = 9)
for(p in 1:ncol(fsib$pheno)) {
  print(colnames(fsib$pheno)[p])
  cofs_fs[[p]] <- cof_selection(fsib, pheno.col = p, k = 2, n.cofactor = 10)
  print(plot(cofs_fs[[p]], main = colnames(fsib$pheno)[p]))
}
dev.off()

for(p in 1:ncol(fsib$pheno)) {
  cofs_fs[[p]][[1]] <- cofs_fs[[p]][[2]] <- cofs_fs[[p]][[3]] <- cofs_fs[[p]][[4]] <- cofs_fs[[p]][[5]] <- NA
}

save(cofs_fs, file = "cofs_fs.RData")
```

```{r}
load("cofs_fs.RData")
```

```{r, eval = FALSE}
for(p in 1:length(cofs_fs)) {
  cofs <- cofs_fs[[p]][[6]][[1]]
  cofs_fs[[p]] <- cof_definition(fsib, pheno.col = p, thres.effect = 0.05, cof.pos = cofs)
}
```

#### Permutation

```{r, eval = FALSE}
set.seed(4321)

cofs_perm <- cim_perm <- vector("list", ncol(fsib$pheno))

for(p in 1:ncol(fsib$pheno)) {
  print(colnames(fsib$pheno)[p])
  cim_perm[[p]] <- cim_scan(fullsib = cofs_fs[[p]], lg = "all", pheno.col = p, 
                            LOD = TRUE, n.perm = 1000, ws = 20)
  print(plot(cim_perm[[p]], main = colnames(fsib$pheno)[p]))
}
dev.off()

save(cim_perm, file = "cim_permutations.RData")
```

```{r}
load("cim_permutations.RData")
```

#### Window size = 20

```{r, eval = FALSE}
cim20 <- vector("list", ncol(fsib$pheno))
pdf("cim20.pdf", width = 16, height = 9)

for(p in 1:ncol(fsib$pheno)) {
  print(colnames(fsib$pheno)[p])
  cim20[[p]] <- cim_scan(fullsib = cofs_fs[[p]], lg = "all", ws = 20, pheno.col = p, LOD = TRUE)
  print(plot(cim20[[p]], main = colnames(fsib$pheno)[p]))
}

dev.off()

save(cim20, file = "cim20.RData")
```

```{r}
load("cim20.RData")
```

#### QTL peaks

```{r}
find_peaks <- function(obj, thr, ws) {
  nlgs <- length(unique(obj[, "lg"]))
  peaks <- vector("list", nlgs)
  peaks1 <- c()
  for (c in 1:nlgs) {
    temp0 <- obj[obj[, "lg"] == c, ]
    if (any(temp0[, "LOD"] > thr)) {
      temp1 <- temp0[temp0[, "LOD"] > thr, ]
      if(is.null(dim(temp1))) {
        peaks0 <- matrix(temp1, nrow = 1, dimnames = list(names(which(temp0[, "LOD"] > thr))))
      } else {
        ranges <- c(0, which(diff(temp1[, "pos.cM"]) > ws), nrow(temp1))
        for (pos in 1:(length(ranges) - 1)) {
          peaks[[c]] <- c(peaks[[c]], names(which.max(temp1[(ranges[pos] + 1):ranges[pos + 1], "LOD"])))
        }
      }
      peaks0 <- temp0[which(rownames(temp0) %in% peaks[[c]]), ]
      peaks1 <- rbind(peaks1, peaks0)
    }
  }
  if(!is.null(peaks1)) {
    rownames(peaks1) <- unlist(peaks) 
  } else {
    peaks1 <- NA
  }
  return(peaks1)
}

peaks <- vector("list", ncol(fsib$pheno))
names(peaks) <- colnames(fsib$pheno)

for(p in c(1:length(peaks))) {
  if(p == 44) next else peaks[[p]] <- find_peaks(obj = cim20[[p]], thr = summary(cim_perm[[p]], alpha = 0.05, verbose = FALSE)[1, 1], ws = 40)
}

knitr::kable(do.call(rbind, peaks))

peaks
```

#### Evaluating QTL phases

```{r, eval = FALSE}
qtl <- segr <- vector("list", ncol(fsib$pheno))
names(qtl) <- names(segr) <- colnames(fsib$pheno)
for(p in c(1:length(qtl))) {
  cat("p =", p, ", trait =", names(qtl)[p], "\n")
  if(!is.null(dim(peaks[[p]]))) {
    qtl[[p]] <- segr[[p]] <- vector("list", nrow(peaks[[p]])) 
    for(q in 1:nrow(peaks[[p]])){ 
      qtl[[p]][[q]] <- cim_char(fullsib = cofs_fs[[p]], pheno.col = p, lg = peaks[[p]][q,"lg"], pos = rownames(peaks[[p]])[q])
      segr[[p]][[q]] <- get_segr(qtl[[p]][[q]])
    }
  }
}

#draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[p]][[q]])
```

Segregation of each QTL:

```{r, eval = FALSE}
#LEAFSHAP1
p <- 16
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[16]][[1]])
get_segr(qtl[[16]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[16]][[2]])
get_segr(qtl[[16]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[16]][[3]])
get_segr(qtl[[16]][[3]])

#LEAFSHAP2
p <- 17
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[17]][[1]])
get_segr(qtl[[17]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[17]][[2]])
get_segr(qtl[[17]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[17]][[3]])
get_segr(qtl[[17]][[3]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[17]][[4]])
get_segr(qtl[[17]][[4]])

#LEAFSHAP3
p <- 18
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[18]][[1]])
get_segr(qtl[[18]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[18]][[2]])
get_segr(qtl[[18]][[2]])

#LEAFSHAP4
p <- 19
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[19]][[1]])
get_segr(qtl[[19]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[19]][[2]])
get_segr(qtl[[19]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[19]][[3]])
get_segr(qtl[[19]][[3]])

#LEFTPP_TP1
p <- 5
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[5]][[1]])
get_segr(qtl[[5]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[5]][[2]])
get_segr(qtl[[5]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[5]][[3]])
get_segr(qtl[[5]][[3]])

#LEFTPP_TP2
p <- 7
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[7]][[1]])
get_segr(qtl[[7]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[7]][[2]])
get_segr(qtl[[7]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[7]][[3]])
get_segr(qtl[[7]][[3]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[7]][[4]])
get_segr(qtl[[7]][[4]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[7]][[5]])
get_segr(qtl[[7]][[5]])

#LEFTPP_TP3
p <- 9
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[9]][[1]])
get_segr(qtl[[9]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[9]][[2]])
get_segr(qtl[[9]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[9]][[3]])
get_segr(qtl[[9]][[3]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[9]][[4]])
get_segr(qtl[[9]][[4]])

#PLANTH_TP1
p <- 11
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[11]][[1]])
get_segr(qtl[[11]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[11]][[2]])
get_segr(qtl[[11]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[11]][[3]])
get_segr(qtl[[11]][[3]])

#PLANTH_TP2
p <- 13
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[13]][[1]])
get_segr(qtl[[13]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[13]][[2]])
get_segr(qtl[[13]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[13]][[3]])
get_segr(qtl[[13]][[3]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[13]][[4]])
get_segr(qtl[[13]][[4]])

#VINDIA
p <- 15
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[15]][[1]])
get_segr(qtl[[15]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[15]][[2]])
get_segr(qtl[[15]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[15]][[3]])
get_segr(qtl[[15]][[3]])

#LA
p <- 30
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[30]][[1]])
get_segr(qtl[[30]][[1]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[30]][[2]])
get_segr(qtl[[30]][[2]])
draw_phase(fullsib = cofs_fs[[p]], fullsib.char = qtl[[30]][[3]])
get_segr(qtl[[30]][[3]])
```

#### Additive and dominance effects of each parent

```{r, eval = FALSE}
#LEAFSHAP1
qtl[[16]]
get_segr(qtl[[16]][[1]])

#LEAFSHAP2
qtl[[17]]

#LEAFSHAP3
qtl[[18]]

#LEAFSHAP4
qtl[[19]]

#LEFTPP_TP1
qtl[[5]]

#LEFTPP_TP2
qtl[[7]]

#LEFTPP_TP3
qtl[[9]] 

#PLANTH_TP1
qtl[[11]]

#PLANTH_TP2
qtl[[13]]

#VINDIA
qtl[[15]]

#LA
qtl[[30]]
```


#### R2 of each QTL

```{r}
#LEAFSHAP1
qtls16 <- r2_ls(fsib, pheno.col = 16, 
                lg = c(3, 10, 12), 
                pos = c("Chr03_2634116", "Chr10_18078811", "Chr12_1378180"))

knitr::kable(qtls16)

#LEAFSHAP2
qtls17 <- r2_ls(fsib, pheno.col = 17, 
                lg = c(3, 10, 11, 12), 
                pos = c("Chr03_2347218", "Chr10_19597490", "Chr11_2819666", "Chr12_3168336"))

knitr::kable(qtls17)

#LEAFSHAP3
qtls18 <- r2_ls(fsib, pheno.col = 18, 
                lg = c(3, 12), 
                pos = c("loc28", "Chr12_2787339"))

knitr::kable(qtls18)

#LEAFSHAP4
qtls19 <- r2_ls(fsib, pheno.col = 19, 
                lg = c(3, 11, 12), 
                pos = c("loc28", "Chr11_2819609", "Chr12_2787339"))

knitr::kable(qtls19)

#LEFTPP_TP1
qtls5 <- r2_ls(fsib, pheno.col = 5, 
               lg = c(3, 12, 13), 
               pos = c("Chr03_7783965", "Chr12_20760521", "Chr13_9041151"))

knitr::kable(qtls5)

#LEFTPP_TP2
qtls7 <- r2_ls(fsib, pheno.col = 7, 
               lg = c(1, 3, 4, 4, 9), 
               pos = c("Chr01_30410817", "Chr03_27381031", "Chr04_3045638", "loc221", "Chr09_22844916"))

knitr::kable(qtls7)

#LEFTPP_TP3
qtls9 <- r2_ls(fsib, pheno.col = 7, 
               lg = c(1, 6, 7, 13), 
               pos = c("Chr01_11083046", "Chr06_23470071", "Chr07_3604826", "Chr13_1910711"))

knitr::kable(qtls9)

#PLANTH_TP1
qtls11 <- r2_ls(fsib, pheno.col = 11, 
                lg = c(9, 10, 12), 
                pos = c("Chr09_1515605", "Chr10_16210815", "Chr12_4116525"))

knitr::kable(qtls11)

#PLANTH_TP2
qtls13 <- r2_ls(fsib, pheno.col = 13, 
                lg = c(1, 3, 3, 4), 
                pos = c("Chr01_29840645", "Chr03_24395656", "Chr03_12011722", "Chr04_29345158"))

knitr::kable(qtls13)

#VINDIA
qtls15 <- r2_ls(fsib, pheno.col = 15, 
                lg = c(2, 14, 15), 
                pos = c("Chr02_8618789", "loc48", "Chr15_2674751"))

knitr::kable(qtls15)

#LA
qtls30 <- r2_ls(fsib, pheno.col = 30, 
                lg = c(7, 11, 12), 
                pos = c("Chr07_1838294", "Chr11_2832259", "Chr12_4356601"))

knitr::kable(qtls30)
```

##### Plot with triangles representing the QTL

```{r, eval = FALSE}
pdf("cim20_ggplot2.pdf", width = 16, height = 9)

for(p in c(1:ncol(fsib$pheno))){
  cim.df <- as.data.frame(cim20[[p]])
  peaks.df <- as.data.frame(peaks[[p]])
  cim_thr <- summary(cim_perm[[p]], alpha = 0.05, verbose = FALSE)[1, 1]
  print(
    ggplot() + 
      geom_line(data = cim.df, aes(x = pos.cM, y = LOD), color = "black") + 
      {if(!is.null(dim(peaks[[p]]))) geom_point(data = peaks.df, aes(x = pos.cM), y = 0, shape = 6, color = "black", size = 2)} + 
      geom_abline(intercept = cim_thr, slope = 0, color = "black") + 
      scale_y_continuous(limits = c(0,31)) +
      labs(title = colnames(fsib$pheno)[p], x = "Position (cM)") + 
      facet_grid(~lg, space = "free_x", scales = "free_x") + scale_x_continuous(breaks = seq(0, 300, 50), labels = seq(0, 300, 50)) + 
      theme_minimal() +
      theme(panel.grid = element_blank())
  )
}

dev.off()
```

##### Better plot

```{r}
names(cim20) <- colnames(fsib$pheno)

ggplot_qtl <- function(cim.df = cim.df, peaks.df = peaks.df, cim_thr = cim_thr){
  cim.df$Phenotype <- factor(cim.df$Phenotype, levels = unique(cim.df$Phenotype))
  peaks.df$Phenotype <- factor(peaks.df$Phenotype, levels = unique(peaks.df$Phenotype))
  cim_thr$Phenotype <- factor(cim_thr$Phenotype, levels = unique(cim_thr$Phenotype))
  peaks.df$y.dat <- -(max(peaks.df$LOD)/50)*rep(1:nlevels(peaks.df$Phenotype), table(peaks.df$Phenotype))
  ggplot() + 
    geom_line(data = cim.df, aes(x = pos.cM, y = LOD, color = Phenotype), alpha = 0.8) + 
    {if(!is.null(dim(peaks.df))) geom_point(data = peaks.df, aes(x = pos.cM, color = Phenotype, y = y.dat), shape = 6, size = 1, stroke = 1, alpha = 0.8)} + 
    geom_hline(data = cim_thr, aes(yintercept = lod, colour = Phenotype), linetype = 2, alpha = 0.8) +
    scale_y_continuous(limits = c(min(peaks.df$y.dat), max(peaks.df$LOD)), labels = scales::label_number(accuracy = 0.1)) +
    labs(x = "Posição (cM)", color = "") +
    # facet_grid(Phenotype ~ lg, space = "free_x", scales = "free_x") +
    facet_grid(. ~ lg, space = "free_x", scales = "free_x", switch = "x") +
    scale_x_continuous(breaks = seq(0, 300, 50), labels = seq(0, 300, 50), expand = expansion(add = 20)) + 
    theme_classic() +
    # theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), # uncomment if you want the numbers
    theme(axis.text.x = element_blank(), axis.title.x = element_blank(), # comment if you want the numbers
          strip.background = element_blank(), strip.placement = "outside", panel.spacing.x = unit(0.2, "lines"),
          legend.position = "bottom", legend.margin = margin(t=-10)) + 
    guides(colour = guide_legend(nrow = 1))
}
```




```{r}
#mudando o nome dos caracteres
names(cim20) <- gsub("PSR_2016", "SRDW_2016", names(cim20))
names(cim20) <- gsub("PFR_2016", "SRFW_2016", names(cim20))
names(cim20) <- gsub("PSV_2016", "VDW_2016", names(cim20))
names(cim20) <- gsub("PFV_2016", "VFW_2016", names(cim20))

names(cim20) <- gsub("PSR_2017", "SRDW_2017", names(cim20))
names(cim20) <- gsub("PFR_2017", "SRFW_2017", names(cim20))
names(cim20) <- gsub("PSV_2017", "VDW_2017", names(cim20))
names(cim20) <- gsub("PFV_2017", "VFW_2017", names(cim20))
                     
names(cim20) <- gsub("LFSIZE_2017", "LFSIZE1_2017", names(cim20))
names(cim20) <- gsub("PETIOL_2017", "PETIOL1_2017", names(cim20))
                     
names(cim20) <- gsub("PFV.plant_2017","VFW_2017", names(cim20))
names(cim20) <- gsub("PSV.plant_2017","VDW_2017", names(cim20))
names(cim20) <- gsub("PFR.plant_2017","SRFW_2017", names(cim20))
names(cim20) <- gsub("PSR.plant_2017","SRDW_2017", names(cim20))
names(cim20) <- gsub("NOSR.plant_2017","NOSR_2017", names(cim20))
```


#### Artigo 1
```{r}
#renomeando o nome dos caracteres só para o artigo
names(cim20) <- gsub("LEAFSHAP1_2016", "LEAFSHAP1", names(cim20))
names(cim20) <- gsub("LEAFSHAP2_2016", "LEAFSHAP2", names(cim20))
names(cim20) <- gsub("LEAFSHAP3_2016", "LEAFSHAP3", names(cim20))
names(cim20) <- gsub("LEAFSHAP4_2016", "LEAFSHAP4", names(cim20))

names(cim20) <- gsub("LEFTPP_TP1_2016", "LEFTPP_TP1", names(cim20))
names(cim20) <- gsub("LEFTPP_TP2_2016", "LEFTPP_TP2", names(cim20))
names(cim20) <- gsub("LEFTPP_TP3_2016", "LEFTPP_TP3", names(cim20))

names(cim20) <- gsub("PLANTH_TP1_2016", "PLANTH_TP1", names(cim20))
names(cim20) <- gsub("PLANTH_TP2_2016", "PLANTH_TP2", names(cim20))

names(cim20) <- gsub("VINDIA_2016", "VINDIA", names(cim20))
names(cim20) <- gsub("LA_2016", "LA", names(cim20))
```

```{r}
#LEAFSHAP1, LEAFSHAP2, LEAFSHAP3, LEAFSHAP4
cim.df <- peaks.df <- cim_thr <- c()
for(p in c(16:19)){
  cim.df <- rbind(cim.df, data.frame(cim20[[p]], Phenotype = names(cim20)[p]))
  {if(!is.null(peaks[[p]])) peaks.df <- rbind(peaks.df, data.frame(peaks[[p]], Phenotype = names(cim20)[p]))}
  cim_thr <- rbind(cim_thr, data.frame(lod=summary(cim_perm[[p]], alpha = 0.05, verbose = FALSE)[1, 1], Phenotype = names(cim20)[p]))
}
A <- ggplot_qtl(cim.df = cim.df, peaks.df = peaks.df, cim_thr = cim_thr); A
ggsave("A.png", width = 10, height = 2.5, units = "in", dpi = 300)


sel.traits <- c("LEFTPP_TP1", "LEFTPP_TP2", "LEFTPP_TP3", "PLANTH_TP1", "PLANTH_TP2", "VINDIA", "LA")
cim.df <- peaks.df <- cim_thr <- c()
for(p in which(names(cim20) %in% sel.traits)){
  cim.df <- rbind(cim.df, data.frame(cim20[[p]], Phenotype = names(cim20)[p]))
  {if(!is.null(peaks[[p]])) peaks.df <- rbind(peaks.df, data.frame(peaks[[p]], Phenotype = names(cim20)[p]))}
  cim_thr <- rbind(cim_thr, data.frame(lod=summary(cim_perm[[p]], alpha = 0.05, verbose = FALSE)[1, 1], Phenotype = names(cim20)[p]))
}
B <- ggplot_qtl(cim.df = cim.df, peaks.df = peaks.df, cim_thr = cim_thr); B
ggsave("B.png", width = 10, height = 2.5, units = "in", dpi = 300)


ggpubr::ggarrange(A, B, labels = c("A", "B"), ncol = 1, nrow = 2)
ggsave("AB.png", width = 8, height = 7, units = "in", dpi = 500)
```


### LinkageMap

```{r}
#install.packages("LinkageMapView")
library(LinkageMapView)
```

```{r}
write_map(chr_lgs_map, "results_onemap.txt")
```

```{r}
sweetpotato <- read.csv("results_onemap.txt", sep = " ", stringsAsFactors = TRUE, header = FALSE)

sweetpotato <- sweetpotato[c(1,3,2)]
colnames(sweetpotato) <- c("Chr", "Pos", "Marker")

sweetpotato$Chr <- as.character(sweetpotato$Chr)
sweetpotato$Pos <- as.numeric(sweetpotato$Pos)
sweetpotato$Marker <- as.character(sweetpotato$Marker)

maxpos <- "2440.47"
at.axis <- seq(0, maxpos)
axlab <- vector()

for (lab in 0:maxpos) {
  if (!lab %% 50) {
    axlab <- c(axlab, lab)
  }
  else {
    axlab <- c(axlab, NA)
  }
}

outfile = file.path(".", "density-map2.pdf")
lmv.linkage.plot(sweetpotato, 
                 outfile, 
                 denmap = TRUE, 
                 cex.axis = 1, 
                 at.axis = at.axis, 
                 labels.axis = axlab, 
                 col.lgtitle = "black", 
                 cex.lgtitle = 1)
```
